---
title: "HW week 12"
subtitle: "w203: Statistics for Data Science"
author: "w203 teaching team"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# OLS Inference

The file videos.txt contains data scraped from Youtube.com.
```{r}
data = read.delim("./videos.txt",header=TRUE , sep="\t")
summary(data)
```

```{r}
hist(data$views)
hist(log(data$views),breaks=50)
hist(log(data$length),breaks=50)
hist(data$rate)

```

1. Fit a linear model predicting the number of views (views), from the length of a video (length) and its average user rating (rate).
```{r}

model1 = lm(log(views)~rate + log(length) , data = data, na.action = na.omit)
summary(model2)
model2 = lm(log(views)~rate + length , data = data, na.action = na.omit)
```


2. Using diagnostic plots, background knowledge, and statistical tests, assess all 6 assumptions of the CLM.  When an assumption is violated, state what response you will take.
```{r}
plot(model1)
plot(model1, pch=".")
plot(model2)
plot(model2, pch=".")

hist(model1$residuals,breaks=50)
hist(model2$residuals, breaks=50)
```


3. Generate a printout of your model coefficients, complete with standard errors that are valid given your diagnostics.  Comment on both the practical and statistical significance of your coefficients.