---
title: "Lab 4: Does Prenatal Care Improve Infant Health?"
author: "w203: Statistics for Data Science"

date: "April 10, 2017"
output: pdf_document
---

# Introduction

This is a group lab.  You may work in teams of 2 or 3.

The file bwght\_w203.RData contains data from the National Center for Health Statistics and from birth certificates.  Your team has been hired by a health advocacy group to study this data and help them understand whether prenatal care improves health outcomes for newborn infants.

The file includes a birthweight variable.  Additonally, the one- and five-minute APGAR scores are included.  These are measures of the well being of infants just after birth.

Variable descriptions are provided as follows.

```{r}
load("bwght_w203.RData")
library(car)
library(lmtest)
desc
head(data)
```

# Assignment

Prepare a report addressing the question of whether prenatal care improves newborn health outcomes.

A successful submission will include

1. A brief introduction

2. A model building process, supported by exploratory analysis.  Your EDA should be interspersed with, and support, your modeling decisions.  In particular, you should use exploratory techniques to address

```{r}
cor(data$bwght,data, use = "complete.obs")
cor(log(data$bwght),data, use = "complete.obs")
cor(log(data$bwght),data, use = "complete.obs")

scatterplotMatrix(data[,],  use = "complete.obs" )

```

```{r}
summary(data)
hist(data$bwght, breaks = 50)
summary(data$omaps)
hist(data$omaps)
table(data$omaps)


summary(data$fmaps)
hist(data$fmaps)
table(data$fmaps)

summary(data$fmaps + data$omaps)
hist(data$fmaps + data$omaps)
table(data$fmaps+ data$omaps)


sapply(colnames(data), function(x) hist(data[,x], breaks = 50, xlab = x ), USE.NAMES= FALSE)
```

```{r}
sapply(colnames(data), function(x){ plot(data$omaps~jitter(data[,x],2), xlab = x ); abline(lm(data$omaps~data[,x]))}, USE.NAMES= FALSE)
```

```{r}
data = data[complete.cases(data),]

summary(data$omaps)
summary(data$npvis)
model1 = lm(data$omaps ~ data$npvis, na.action = na.omit)

summary(model1)
coeftest(model1, vcov = vcovHC)

pa1r(mfrow=c(2,2))
plot(model1)

## Assumption 1
## THe model is linear in coefficients by design

## Assuption 2
##Since the data is  from National Center for Health Statistics we can say it is a reliable source, satisfying assumption 2

## Assumption 3
## Call since there is only one term there is no way for co-linearity among variables

## Assumption 4
## The plot residual vs fit plot is centered around 0 so we can satisfy assumption 4



## Assumption 5
## for Homoskedasitic test
bptest(model1)

## We fail to reject the null hypothesis hence the model has Homoskedasitic
## We will be using Homoskedasitic roboust coef

## Assumptioin 6
## Since we have a lot of data point we can assumption asymptotics of OLS co-efficients

```

```{r}
head(data[data$omaps==0,])

model2 = lm(data$omaps ~ data$npvis + log(data$bwght), na.action = na.omit)

coeftest(model2, vcov. = vcovHC)
summary(model2)


model11 = lm(data$omaps ~ data$npvis + (data$bwght), na.action = na.omit)

coeftest(model11, vcov. = vcovHC)
summary(model11)

model7 = lm(log(data$omaps) ~ data$npvis + log(data$bwght), na.action = na.omit)

coeftest(model7, vcov. = vcovHC)
summary(model7)

model8 = lm(log(data$omaps) ~ data$npvis + log(data$bwght) + data$monpre, na.action = na.omit)

coeftest(model8, vcov. = vcovHC)
summary(model8)


?waldtest(model8,model2, vcov= vcovHC)
summary(model1)
```



```{r}

summary(data$monpre)

data$monpre = data$monpre+1
summary(data$monpre)

model3 = lm(data$omaps ~ data$npvis + data$mage + data$meduc + data$monpre + data$fage + data$feduc + data$bwght, na.action = na.omit)
plot(model3)
vif(model3)

coeftest(model3, vcov = vcovHC)
summary(model3)

model4 = lm(data$omaps ~ data$npvis +data$fage+ data$mage + (data$fage - data$mage)^2 +data$meduc +data$monpre + ( data$feduc) + log(data$bwght), na.action = na.omit)

vif(model4)
coeftest(model4, vcov = vcovHC)
summary(model4)

model5 = lm(data$omaps ~ data$npvis + data$fage + data$mage +data$meduc +data$monpre + data$feduc + log(data$bwght), na.action = na.omit)
coeftest(model5, vcov = vcovHC)
summary(model5)

model9 = lm(data$omaps ~ data$npvis + data$fage + data$mage*data$cigs +data$cigs +data$meduc +data$monpre + data$feduc + log(data$bwght), na.action = na.omit)
coeftest(model9, vcov = vcovHC)
summary(model9)

model10 = lm(data$omaps ~ data$npvis + data$mage*data$drink +data$drink  +data$monpre  + log(data$bwght), na.action = na.omit)
coeftest(model10, vcov = vcovHC)
summary(model10)


data$omaps = data$omaps+1

model6 = lm(log(data$omaps) ~ data$npvis + log(data$fage + data$mage) +data$meduc +data$monpre + data$feduc + log(data$bwght), na.action = na.omit)
coeftest(model6, vcov = vcovHC)
summary(model6)

waldtest(model6,model7,vcov = vcovHC )

plot(model6)
```


- What transformations to apply to variables and what new variables should be created.


- What variables should be included in each model

- Whether model assumptions are met

3. A minimum of three model specifications.  In particular, you should include

- One model with only the explanatory variables of key interest.

- One model that includes only covariates that you believe increase the accuracy of your results without introducing bias.

- One model that includes the previous covariates, but also covariates that may be problematic for one reason or another.

4. For your first model, a detailed assessment of the 6 CLM assumptions.  For additional models, you should check all assumptions, but only highlight major differences from your first model in your report.

5. A well-formatted regression table summarizing your model results.  Make sure that standard errors presented in this table are valid.  Also be sure to comment on both statistical and practical significance.

6. A discussion of whether your results can be interpretted causally.  In particular, include a discussion of what variables are not included in your analysis and the likely direction of omitted variable bias.  Also include a discussion of which included variables may bias your results by absorbing some of the causal effect of prenatal care.

7. A brief conclusion with a few high-level takeaways.


Please limit all submissions to 30 pages.  Be sure to turn in both your pdf report and also your source code.
